\usefont{OT1}{bch}{b}{n} \normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt] \huge Deep learning transcriptomic model for prediction of pan-drug chemotherapeutic sensitivity \\

### 2020-06-23 manuscript/manuscript.tex
updated figures

% Abstract
\begin{abstract}
	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles. This development would allow for identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden. As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model. A cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups. Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity. Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value. The feature space was reduced to 300 genes, which functional profiling indicated was primarily enriched for the focal adhesion, ECM-receptor and proteoglycan interaction pathways. Using these selected genes, a deep learning neural network architecture was developed to predict response groups. It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparamter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).  This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs. \vspace{1em}

@@

% clustering figure
\begin{figure*}[!ht]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=2.5in]{Figures/pca.png}
        \caption{}
        \label{fig:pca}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=2.5in]{Figures/volcano.png}
        \caption{}
        \label{fig:volcano}
    \end{subfigure}

>>>>

% Abstract
\begin{abstract}
	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles. This development would allow for identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden. As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model. A cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups. Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity. Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value. The feature space was reduced to 300 genes; functional profiling was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways. Using these selected genes, a deep learning neural network architecture was developed to predict response groups. It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparameter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).  This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs. \vspace{1em}

@@

% clustering figure
\begin{figure*}[!ht]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=2.6in]{Figures/pca.png}
        \caption{}
        \label{fig:pca}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=2.5in]{Figures/volcano.png}
        \caption{}
        \label{fig:volcano}
    \end{subfigure}
